InnovaChallengeBBVA
========================================================
Downloading and saving the data from the BBVA API
--------------------------------------------------------

#### 1. Set the working directory and load the required libraries.
```{r wd_libraries}
setwd("~/Dropbox/R InnovaChallengeBBVA")
require(RCurl)
require(rjson)
```


#### 2. Set the API authentication parameters.
```{r authentication_parameters}
APP_ID = "martinezBBVA"
APP_KEY = "fd48d0216bcdba771d7db97c4578c05d6a043428"
strToEncode = paste(APP_ID,APP_KEY,sep=":")
auth = base64Encode(strToEncode)  # in RCurl package
```

#### 3. Get the metadata that will be needed from the API.   

> The categories' codes from the API. 

```{r categories}
options(stringsAsFactors = FALSE)
CATEGORIES <- "https://api.bbva.com/apidatos/info/merchants_categories.json"
cat <-getForm(CATEGORIES, .opts=list(userpwd=strToEncode, httpauth=1L,verbose=T))
cat <- fromJSON(cat)
categories <- do.call(rbind.data.frame, cat$data$categories)
names(categories) <- names(cat$data$categories[[1]])

# Add a row with the 'all' category
categories <- rbind(categories, c("all", "All categories"))

write.csv(categories, "metadata/merchants_categories.csv", row.names = FALSE)
```

> The postal codes list  

```{r postal_codes}
postal_codes <- read.table("~/Dropbox/R Data/cps_madrid.csv", sep="|", header=T, stringsAsFactors=F)

write.csv(postal_codes, "metadata/postal_codes.csv", row.names = FALSE)
```
   
> The dates.   

```{r dates}
dates <- c("201211", "201212", "201301", "201302", "201303", "201304")
write.csv(dates, "metadata/dates.csv", row.names = FALSE)
```


#### 4. Download top client origin zip code.   

> CP destiny: every CP.     
> CATEGORIES: ALL and every category.   
> DATA: CP origin.    
> DATE: every month.    

Load the metadata needed to run the loop and set the request parameters
```{r loop_and_request_parameters}
# Read the metadata.
dates <- read.csv("metadata/dates.csv", stringsAsFactors = FALSE)
postal_codes <- read.csv("metadata/postal_codes.csv", stringsAsFactors = FALSE)
categories <- read.csv("metadata/merchants_categories.csv", stringsAsFactors = FALSE) 

# API url
CUSTOMER_ZIPCODES="https://api.bbva.com/apidatos/zones/customer_zipcodes.json"

# Set the requested parameters
datemin = 20121101
datemax = 20130430
groupby = "month"
by = "incomes"
```

> Downlaod the data from the API

```{r loop}
# LOOP for downloading the data (all CP, all and every category, every month)

zipcodes <- data.frame(
    label = character(),
    num_cards = integer(),
    num_payments = integer(),
    incomes = numeric(),
    stringsAsFactors=F)

for (category in categories$code) {
	for (zipcode in postal_codes$label) {
		origin <- tryCatch(getForm(CUSTOMER_ZIPCODES, 
								   .opts=list(userpwd=strToEncode, httpauth=1L,verbose=T), 
								   date_min = datemin,date_max = datemax, group_by = groupby, 
								   zipcode = zipcode, category = category, by=by), 
						   HTTPError = function(e) { 
						   	cat("HTTP error: ", e$message, "\n")
						   	})
		
		if (length(origin != 0)) {
			origin <- fromJSON(origin)
			for (h in (1:length(origin$data$stats))) {
				
				if (length(origin$data$stats[[h]]$zipcodes) != 0) {
						df <- do.call(rbind.data.frame, origin$data$stats[[h]]$zipcodes)
						df$month <- origin$data$stats[[h]]$date
						df$cp_dest <- as.character(zipcode)
						df$category <- category
						zipcodes <- rbind(zipcodes, df)
				}
			}
		}
	}
}
```

Add a row for any category-cp-month where there is no label (origin) == cp_dest
```{r add_rows}
temp2 <- data.frame(
	label = character(), 
	num_cards = integer(), 
	num_payments = integer(), 
	incomes = numeric(), 
	month = character(), 
	cp_dest = character(),
	category = character()
	)
    i = 1

for (cat in unique(zipcodes$category)){
	temp <- zipcodes[zipcodes$category == cat, ]
	for (cp in unique(zipcodes$cp_dest)){
		temp <- temp[temp$cp_dest == cp, ]
		for (date in unique(zipcodes$month)){
			a <- temp[temp$month == date & temp$cp_dest == temp$label, ]
			if(length(a[ ,1]) == 0){
				temp2[i, ] <- c(cp, 0, 0, 0, date, cp, cat)
				i <- i+1
			}
		}
	}
}
```

Combine both the 'zipcodes' and the 'temp2' data.frames
```{r combine_zipcodes_temp2}
zipcodes <- rbind(zipcodes, temp2)

# Replace the 'label' variable name for 'cp_origin'
colnames(zipcodes)[names(zipcodes) == "label"] <- "cp_origin"


# Save it as a csv file
write.csv(zipcodes, "rawdata/all_zipcodes.csv", row.names = FALSE)
```
