InnovaChallengeBBVA
========================================================
Load the needed shapefiles
--------------------------------------------------------

```{r libraries}
setwd("/Users/martinez/Dropbox/R InnovaChallengeBBVA")
library(ggmap)
library(maptools)
library(sp)
library(rgdal)
library(rgeos)
library(ggplot2)
```

#### 1. Load the postal_codes polygons

```{r load_postal_codes}
# Read the shapefile
getinfo.shape("Shapefiles/MAD_Codigos_Postales/codigos_postales.shp")
cp_poly <- readShapePoly("Shapefiles/MAD_Codigos_Postales/codigos_postales.shp", proj4string = ED50)

# Change the encoding
cp_poly@data <- as.data.frame(apply(cp_poly@data, 2, function(x) iconv(x, "latin1", "UTF-8")))

# Set the projection to long lat
cp_poly <- spTransform(cp_poly, CRS("+proj=longlat"))

# Subset the Madrid capital postal codes
cp_poly_mad <- cp_poly[grep("^280[0-9]{2}", cp_poly$GEOCODIGO), ]

rm(cp_poly)
```


#### 2. Get the Shopping Centers' locations

> load the locations (for zara and el corte inglés) 

```{r load_comercios}
comercios <- read.csv("metadata/comercios.csv", header = TRUE, stringsAsFactors = FALSE)

# Based on the address, get the long lat for every shop.
comercios_lonlat <- sapply(comercios$Location, geocode)
comercios_lonlat <- as.data.frame(t(comercios_lonlat))

comercios_lonlat$Location <- row.names(comercios_lonlat)
comercios_lonlat <- merge(comercios, comercios_lonlat)
head(comercios_lonlat)
```


> Get the long lat for the 'centros comerciales' and the 'Grandes superficies especializadas' from the Comunidad de Madrid Shapefiles

```{r cc_gse_shapefiles}

# Set the shapefiles original projection
ED50<-CRS(paste("+proj=utm +zone=30 +ellps=intl +units=m +no_defs"))

# Read the two Shapefiles
centros_comerciales <- readShapePoints("Shapefiles/MAD_Centros comerciales/centrocom.shp",  proj4string = ED50)

grandes_superficies <- readShapePoints("Shapefiles/MAD_Grandes superficies especializadas/comerespecial.shp", proj4string = ED50)

#Add the variable 'TIPO' to the grandes_superficies@data
grandes_superficies$TIPO <- "grande superficie"

# Bind the two shapefiles
cc <- spRbind(centros_comerciales, grandes_superficies)

# Set the right enconding
cc@data <- as.data.frame(apply(cc@data, 2, function(x) iconv(x, "latin1", "UTF-8")))

# Subset the ones in Madrid Capital
cc_mad <- cc[cc$MUNICIPIO == "Madrid", ]

# Set the projection to long lat
ccLongLat <- spTransform(cc_mad, CRS("+proj=longlat"))
```

> Put the three together

```{r}
# Transform the 'comercios_lonlat' data.frame into a SpatialPointsDataFrame

comerciosSP <- SpatialPointsDataFrame(comercios_lonlat[ ,5:6], proj4string = CRS("+proj=longlat"), data = comercios_lonlat[ ,1:4])

# Get ready the comerciosSP@data to bind it to ccLongLat
names(comerciosSP@data)
names(ccLongLat@data)

comerciosSP$Location <- NULL
comerciosSP$OBS <- NULL
comerciosSP$CMUN <- "079"
comerciosSP$MUNICIPIO <- "Madrid"
colnames(comerciosSP@data)[names(comerciosSP@data) == "Via"] <- "ETIQUETA"
colnames(comerciosSP@data)[names(comerciosSP@data) == "Text"] <- "TIPO"

comerciosSP@data <- comerciosSP@data[ ,c(3,4,2,1)]

names(comerciosSP@data)
names(ccLongLat@data)

# Bind the SpatialPintsDataFrames both 'comerciosSP' and 'ccLongLat'
comerciales <- spRbind(comerciosSP, ccLongLat)

comerciales <- comerciales[grepl("^El Corte Ingl", comerciales$ETIQUETA) == FALSE, ]

# Transform them into a data frame
long <- coordinates(comerciales)[, 1]  
lat <- coordinates(comerciales)[, 2]
dataComerciales <- cbind(comerciales@data, long, lat)
head(dataComerciales)
```


#### 3. Load the metro stations
```{r metro_stations}
# Load the shapefile
getinfo.shape("Shapefiles/MAD_metro_stations/estmetro.shp")
metroStations <- readShapePoints("Shapefiles/MAD_metro_stations/estmetro.shp", proj4string = ED50)

# Change the encoding
metroStations@data <- as.data.frame(apply(metroStations@data, 2, function(x) iconv(x, "latin1", "UTF-8")))

# Set the projection to long lat
metroStations <- spTransform(metroStations, CRS("+proj=longlat"))

# Subset the metroStations points in the cpPolyMad polygons
over(metroStations,as(cp_poly_mad,"SpatialPolygons"))
metroStationsMad <- metroStations[!is.na(over(metroStations,as(cp_poly_mad,"SpatialPolygons"))), ]

# Subset the main metro stations (just to have a reference)
main <- c("Delicias", "Hospital 12 de Octubre", "Pitis", "Mirasierra", "Montecarmelo", "Fuencarral", "Barrio del Pilar", "Chamartin", "Plaza de Castilla", "Francos Rodriguez", "Las Tablas", "Pinar de Chamartín", "Mar de Cristal", "Santiago Bernabéu", "Oporto", "Nuevos  Ministerios", "Ciudad Lineal", "Vicálvaro", "Sáinz de Baranda", "Ventas", "O'Donnell", "Barrio de la Concepción", "Alonso Martínez", "Embajadores", "Sol", "Moncloa", "Plaza de España", "Aeropuerto T4", "Aeropuerto", "Campo de las Naciones", "Arturo Soria", "Carabanchel Alto", "Carabanchel", "Villaverde Bajo Cruce", "Villaverde Alto", "La Peseta", "Las suertes", "Puente de Vallecas", "Villa de Vallecas", "Príncipe Pío", "Lago", "Casa de Campo", "Aluche", "Cuatro Vientos", "La Peseta", "San Bernardo", "Ciudad universitaria", "Puerta de Toledo", "Estación de Aravaca", "Antonio Machado", "Cuatro Caminos", "Opera", "Ibiza", "Goya", "Avda. de América", "Pacífico", "García Noblejas", "Canillejas", "Estadio Olímpico", "La Elipa", "Puerta de Arganda", "Alto del Arenal", "Vinateros", "Las Suertes", "Atocha", "Prosperidad", "Colombia")

metroStationsMad <- metroStationsMad[metroStationsMad$ESTACION %in% main, ] 

# transform into a data.frame
long <- coordinates(metroStationsMad)[, 1]  
lat <- coordinates(metroStationsMad)[, 2]
dataMetroStations <- cbind(metroStationsMad@data, long, lat)
head(dataMetroStations)
```

#### 4. Define a function to find out the contiguous polygons 
```{r contiguos_polygons}

cp_poly_mad$GEOCODIGO <- as.numeric(as.character(cp_poly_mad$GEOCODIGO))


getContiguous <- function(mainCP){
	contiguous <- data.frame()
	mainCP <- as.numeric(mainCP)
	for (cp in postal_codes$label){
		print(mainCP)
		print(cp)
		intersec <- gIntersection(as(cp_poly_mad[cp_poly_mad$GEOCODIGO == mainCP, ], "SpatialPolygons"), as(cp_poly_mad[cp_poly_mad$GEOCODIGO == cp, ], "SpatialPolygons"))
		if(length(intersec) != 0 & cp != mainCP){
			tempContiguos <- data.frame(
				"mainCP" = mainCP,
				"contiguousCP" = cp)
			contiguous <- rbind(contiguous, tempContiguos)
			}
		}
	return(contiguous)
	}
```
